---
title: "Assignment 3"
author: "Anjali Pal"
date: "07/05/2021"
output: html_document
---
Please use data set COVID19.csv for next questions.  This data set is in a raw format. You have to clean this dataset before any analysis. Data set is totally raw downloaded from worldmeters today (March 1st 2021)  (https://www.worldometers.info/coronavirus/). Hence, data cleaning and manipulation are required. Please also explain your results with numerical summary as well as graphs, wherever it is applicable. Please check your data also, if it is skewed or normally distributed. Results should be based on skewness or normal distribution. 

The COVID19.csv data set is raw data extracted from Worldometer (March 1st, 2021). Use COVID19.csv data to answer the questions considering the following:

1. The data set requires data cleaning and manipulation before any analysis.
2. Explain your results with a numerical summary and graphs, wherever it is         applicable.
3. The results should be based on skewness or normal distribution, therefore, do    check your data if it is skewed or normally distributed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning=FALSE}
covid_data <- read.csv('C:/Users/Lenovo PC/Downloads/COVID19.csv')
str(covid_data)
```
# Step 1
Keep rows containing country information and remove the rest of the rows. Apart from the country, you should not be having any other row. Move on to the next step only after finishing this.
```{r message = FALSE, warning=FALSE}
covid_data$Country.Other
#First 8 and Last 8 entries are either empty, continents, world or Total
covid_data <- covid_data[9:229,]
unique(covid_data$Country.Other)
```
# Step 2
Calculate the missing percentage of each column using a function. If any column has missing data more than 5%, please remove it. (Do not try this for rows)
```{r message = FALSE, warning=FALSE}
covid_data<-data.frame(lapply(covid_data, gsub, pattern = ",|\\+", replacement = ""))
covid_data<-data.frame(lapply(covid_data, gsub, pattern = "N/A", replacement = NA))
colSums(is.na(covid_data)/nrow(covid_data))*100
#Only 7 is above 5 
covid_data <- covid_data[,c(-2,-7)]
```
#Step 3
Give a better column name after cleaning your data.
```{r message = FALSE, warning=FALSE}
colnames(covid_data)[colnames(covid_data) %in% 
                       c("Country.Other", "TotalCases","NewCases","TotalDeaths","TotalRecovered","NewRecovered","ActiveCases","Serious.Critical","Tot.Cases.1M.pop", "Deaths.1M.pop","TotalTests", "Tests.1M.pop", "Population", "Continent","X1.Caseevery.X.ppl", "X1.Deathevery.X.ppl", "X1.Testevery.X.ppl")]<- c("Country", "Total_Cases", "New_Cases","Total_Deaths","Total_Recovered","New_Recovered","Active_Cases","Critical_Cases","Total_Cases_pm", "Deaths_pm","Total_Tests", "Tests_pm", "Population", "Continent","Case_Every_X_ppl", "Death_Every_X_ppl", "Test_Every_X_ppl")
colnames(covid_data)
```
#Converting to numeric
```{r message = FALSE, warning=FALSE}
str(covid_data)
library(dplyr)
covid_data <- covid_data %>%mutate_at(c(1,3:14,16:18), as.numeric,na.rm = TRUE)
str(covid_data)
```
Create plots for total cases, total death, and total recovery. Explain with a figure for each.
```{r message = FALSE, warning=FALSE}
hist(covid_data$Total_Cases)
library(ggplot2)
ggplot(covid_data, aes(x =Total_Cases))+geom_histogram(aes(y= ..density..),colour = "black",fill = "white",binwidth = 5)+stat_function(fun= dnorm,colour = "red",args = list(mean = mean(covid_data$Total_Cases), sd = sd(covid_data$Total_Cases)))
ggplot(covid_data, aes(x =Total_Deaths))+geom_histogram(aes(y= ..density..),colour = "black",fill = "white",binwidth = 5)+stat_function(fun= dnorm,colour = "red",args = list(mean = mean(covid_data$Total_Deaths), sd = sd(covid_data$Total_Deaths)))
ggplot(covid_data, aes(x =Total_Recovered))+geom_histogram(aes(y= ..density..),colour = "black",fill = "white",binwidth = 5)+stat_function(fun= dnorm,colour = "red",args = list(mean = mean(covid_data$Total_Recovered,na.rm = TRUE), sd = sd(covid_data$Total_Recovered,na.rm = TRUE)))

```
Create a plot to examine the correlation between total cases and total population. Explain if there is any correlation between total cases and total population.
```{r message = FALSE, warning=FALSE}
plot(covid_data$Total_Cases,covid_data$Population)
cor(covid_data$Total_Cases,covid_data$Population,method="pearson")
```
Create a plot to examine the correlation between Tot Cases/1M pop and total population. Explain if there is any correlation between them?
```{r message = FALSE, warning=FALSE}
plot(covid_data$Total_Cases_pm,covid_data$Population)
cor(covid_data$Total_Cases_pm,covid_data$Population,method="pearson")
```
Which column do you feel is better for comparison purposes, total cases or TotCases/1M pop. Explain.
```{r message = FALSE, warning=FALSE}
#Tot Cases per million as if population is less, cases are less. But it doesn't mean that it's doing better than other nations with more population.
```
Create a plot to examine the correlation between total cases and total death. Explain the figure.
```{r message = FALSE, warning=FALSE}
plot(covid_data$Total_Cases,covid_data$Total_Deaths)
cor(covid_data$Total_Cases,covid_data$Total_Deaths,method="pearson")
```
Create a plot to examine the correlation between total cases and Deaths/1M pop. Explain the figure. Which column is more suitable to compare the result, total death or Death/1Mpop?
```{r message = FALSE, warning=FALSE}
plot(covid_data$Total_Cases,covid_data$Deaths_pm)
cor(covid_data$Total_Cases,covid_data$Deaths_pm,method="pearson")
#Deaths PM , better comparison.
```
Compare Tot Cases/1M pop by continent, and explain your result. 
```{r message = FALSE, warning=FALSE}
levels(covid_data$Continent)
boxplot(Total_Cases_pm ~ Continent, data = covid_data,na.exclude(covid_data$Continent),main = "Total Cases/1M pop by continent",col = "grey", frame = FALSE)
#Europe has highest total number of cases per million population on average
#Australia/Oceania has the least on average
```
Compare Deaths/1M pop by continent, and explain your result. 
```{r message = FALSE, warning=FALSE}
boxplot(Deaths_pm ~ Continent, data = covid_data,na.exclude(covid_data$Continent),main = "Total deaths/1M pop by continent",col = "grey", frame = FALSE)
#Europe has highest death cases per million population on average
#Australia/Oceania has the least on average
```
Which country is best among testing the COVID19 and which country is worst? There are two columns total test vs. test/M. Choose appropriate column. 
```{r message = FALSE, warning=FALSE}
covid_data$Country[which.max(covid_data$Tests_pm)]
#ARUBA is the best
covid_data$Country[which.min(covid_data$Tests_pm)]
#ALGERIA IS WORST TESTING
```
Compare your COVID19 test results by continent? There are two columns total test vs test/M. Choose appropriate column. 
```{r message = FALSE, warning=FALSE}
ggplot(data=covid_data, aes(x=Continent, y=Tests_pm),frame=TRUE) +
  geom_bar(stat="identity", fill="pink")
```
Check if Tests/1M pop is skewed or normally distributed.
```{r message = FALSE, warning=FALSE}
shapiro.test(covid_data$Tests_pm)
#Pvalue is 1.284e-07 <0.05, Ho may be rejected. It is not normally distributed.
qqnorm(covid_data$Tests_pm)
qqline(covid_data$Tests_pm)
#Data is clearly skewed.
```